---
title: "Establishment Entry Rate"
subtitle: "Analyzing the Impact"
format: html
execute:
    echo: false
    warning: false
    message: false
---

```{r}
library(tidyverse)
library(tidycensus)
library(janitor)
```

```{r}
bds2022 <- read_csv("bds2022_msa.csv") %>%
  clean_names() %>%
  filter(year == 2022) %>%
  select(
    msa,
    estabs_entry,        
    estabs_exit,  
    estabs_entry_rate,  # startup rate
    estabs_exit_rate,   # business exit rate
    firms, estabs, emp,
    firmdeath_estabs,
    job_creation_rate, 
    net_job_creation_rate,
    job_creation_rate_births
  ) %>%
  drop_na()
```

```{r}
# Load crosswalk file
# Load dataset with metropolitan area names matching to codes
msa_names <- read_csv("list1_2020.csv") %>%
  clean_names() %>%
  transmute(
    cbsa_code = as.character(cbsa_code),
    msa_name = cbsa_title
  )

# Clean BDS data and convert code to character
bds2022 <- bds2022 %>%
  mutate(msa = as.character(msa)) %>%
  left_join(msa_names, by = c("msa" = "cbsa_code")) %>%
  select(msa_name, everything())
```

# Exit Rate

## Plot 1

```{r}
bds2022 |>
  distinct() |> 
ggplot(aes(x = estabs_entry_rate, y = estabs_exit_rate, label = msa_name)) +
  geom_point(color = "black", fill = "black", alpha = 0.2, size = 3) +
  ggrepel::geom_text_repel(data = subset(bds2022, estabs_entry_rate > 20)) +
  geom_smooth(method = "lm", color = "skyblue", se = FALSE) + 
  scale_x_continuous(limits = c(5, 25)) + 
  scale_y_continuous(limits = c(5, 20)) + 
  labs(
    title = "2022 Metropolitan Area Establishment Entry Rates vs Exit Rates",
    subtitle = "As entry rate increases, exit rate also increases",
    x = "Entry Rate (%) (Establishment Entries per 100 Existing)",
    y = "Exit Rate (%) (Establishment Exits per 100 Existing)",
    caption = "Source: U.S. Census Bureau BDS 2022"
  ) +
  theme_minimal()
```

Understanding economic growth and how to manipulate it becomes increasingly important as societies aim to develop. Using data from the US Census Bureau, I aim to explore the effect of establishment entry rates to establishment exit rates and job growths in different areas of the US in 2022. There is a moderately strong linear correlation between establishment entry rate and exit rate, meaning that a higher entry rate will correlate with a higher exit rate. Thus, despite having a high establishment entry rate that would indicate economic growth, establishments would also increasingly leave that economy. One way to explain this is older shops closing down, and new shops taking their place.

```{r}
library(tidyverse)

plot_data_1 <- bds2022 %>%
  distinct(estabs_entry_rate, job_creation_rate, .keep_all = TRUE)

p1 <- ggplot(plot_data_1, aes(x = estabs_entry_rate, y = job_creation_rate)) +
  geom_point(alpha = 0.2, size = 3) +
  geom_smooth(method = "lm", color = "skyblue", se = FALSE) +
  theme_minimal() +
  labs(title = "Job Creation Rate vs Establishment Entry Rate",
       x = "Entry Rate (%) (Establishment Entries per 100 Existing)",
       y = "Job Creation Rate (%)")

# Remove duplicate points for job_creation_rate_births vs net_job_creation_rate
plot_data <- bds2022 %>%
  distinct(estabs_entry_rate, job_creation_rate_births, .keep_all = TRUE)

# Scatter plot with regression line
p2 <- ggplot(plot_data, aes(x = estabs_entry_rate, y = job_creation_rate_births)) +
  geom_point(alpha = 0.2, size = 3) +
  geom_smooth(method = "lm", color = "skyblue", se = FALSE) +
  theme_minimal() +
  labs(
    title = "Establishment Entry Rate vs Job Creation Rate (Births)",
    x = "Entry Rate (%) (Establishment Entries per 100 Existing)",
    y = "Job Creation Rate From New Establishments (%)"
  )
```

# Job Impacts

## Plot 2
```{r}
# Display plots
p1
```

A strong linear trend is observed between establishment entry rate and job creation rate. As entry rate increases, the overall job creation rate of the area also increases. Sheridan, WY, the right-most point, has a job creation rate slightly below than what the model predicts, but not too far off.

## Plot 3
```{r}
# Display plots
p2
```

Once again, a very stron glinear trend is observed between entry rate and job creation rate, but this time the job creation rate is solely from those new establishments. As entry rate increases, it is self explanatory why the new job creation rate also increases.