---
title: "Mathematical Models"
format: html
execute:
    echo: false
    message: false
    warning: false
---

```{r}
library(tidyverse)
library(tidycensus)
library(janitor)
```

```{r}
bds2022 <- read_csv("bds2022_msa.csv") %>%
  clean_names() %>%
  filter(year == 2022) %>%
  select(
    msa,
    estabs_entry,        
    estabs_exit,  
    estabs_entry_rate,  # startup rate
    estabs_exit_rate,   # business exit rate
    firms, estabs, emp,
    firmdeath_estabs,
    job_creation_rate, 
    net_job_creation_rate,
    job_creation_rate_births
  ) %>%
  drop_na()
```

## Table

```{r}
library(broom)
library(dplyr)
library(knitr)

# Fit models
model1 <- lm(estabs_exit_rate ~ estabs_entry_rate, data = bds2022)
model2 <- lm(net_job_creation_rate ~ estabs_entry_rate, data = bds2022)
model3 <- lm(job_creation_rate_births ~ estabs_entry_rate, data = bds2022)

# Function to prettify term names
pretty_terms <- function(term) {
  case_when(
    term == "(Intercept)" ~ "Intercept",
    term == "estabs_entry_rate" ~ "Establishment Entry Rate",
    TRUE ~ term
  )
}

# Extract tidy summaries with nicer model titles
tidy_m1 <- broom::tidy(model1) %>%
  mutate(
    Model = "Plot 1: Establishment Exit Rate ~ Establishment Entry Rate",
    Term = pretty_terms(term)
  ) %>%
  select(Model, Term, estimate, std.error, statistic, p.value)

tidy_m2 <- broom::tidy(model2) %>%
  mutate(
    Model = "Plot 2: Net Job Creation Rate ~ Establishment Entry Rate",
    Term = pretty_terms(term)
  ) %>%
  select(Model, Term, estimate, std.error, statistic, p.value)

tidy_m3 <- broom::tidy(model3) %>%
  mutate(
    Model = "Plot 3: Job Creation Rate (Births) ~ Establishment Entry Rate",
    Term = pretty_terms(term)
  ) %>%
  select(Model, Term, estimate, std.error, statistic, p.value)

# Combine results
all_models <- bind_rows(tidy_m1, tidy_m2, tidy_m3)

# Rename columns for display
colnames(all_models) <- c("Model", "Term", "Estimate", "Std. Error", "t value", "p value")

# Display table
kable(all_models, digits = 4, caption = "Regression Model Coefficients")
```

## Models

**Model 1** (for plot 1)
$$
\text{estabs\_exit\_rate}_i = \beta_0 + \beta_1 \cdot \text{estabs\_entry\_rate}_i + \epsilon_i
$$

**Model 2** (for plot 2)
$$
\text{net\_job\_creation\_rate}_i = \beta_0 + \beta_1 \cdot \text{estabs\_entry\_rate}_i + \epsilon_i
$$

**Model 3** (for plot 3)
$$
\text{job\_creation\_rate\_births}_i = \beta_0 + \beta_1 \cdot \text{estabs\_entry\_rate}_i + \epsilon_i
$$


The first model examines the relationship between the establishment entry rate and the establishment entry rate. It is specified as a simple linear regression where the exit rate for observation ð‘– is modeled as a function of a constant intercept term  $\beta_0$, the establishment entry rate multiplied by its slope coefficient $\beta_1$, and a random error term $\epsilon_i$, capturing unexplained variation. This model quantifies how changes in the rate at which new establishments enter the market are associated with changes in the rate at which existing establishment leave the market.

The second and third models similarly use a linear regression framework to relate the net job creation rate and job creation rate from births to the establishment entry rate. Here, for both, the dependent variable is the job creation rate attributed specifically to new business births, modeled as a linear function of the establishment entry rate, an intercept, and a random error term. All models assume linearity and normally distributed residuals, providing a basis to estimate and test the statistical significance of the association between business entry dynamics and employment growth.
 